---
output:
  html_document:
    title: "COVIZ: COVID_19 in Latin America"
    author: "Brandon Janes"
    date: "3/22/2020"
    theme: cerulean
    highlight: haddock
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: no
    toc_depth: '3'
---
# COVIZ: COVID-19 in Latin America
By Brandon Janes
```{r import libraries,message=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(data.table)
```
# **Upload Data**
Data from repository for the 2019 Novel Coronavirus Visual Dashboard operated by the [**Johns Hopkins University Center for Systems Science and Engineering**](https://github.com/CSSEGISandData/COVID-19) (JHU CSSE). Also, countries and continents from [**dbouquin**](https://github.com/dbouquin/IS_608/blob/master/NanosatDB_munging/Countries-Continents.csv) and population from [**Gapminder**](gapminder.org).
```{r, import data}
confirmed <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", stringsAsFactors = FALSE)
region <- read.csv("https://raw.githubusercontent.com/dbouquin/IS_608/master/NanosatDB_munging/Countries-Continents.csv", stringsAsFactors = FALSE)
pop <- read.csv("population_total.csv", stringsAsFactors = FALSE) # thanks Hans!
```
```{r data wrangling, include=FALSE}
confirm <- select(confirmed, -(Lat:Long))
matriz <- select(confirm, -(Province.State))
matriz[-1] = as.numeric(unlist(matriz[-1]))
matriz[is.na(matriz)] <- 0
# group by
df <- matriz %>% group_by(Country.Region) %>% summarise_all(funs(sum))
# transpose
transpose_df <- as.data.frame(t(df))
# make country names column headers with custom function
header.true <- function(df) {
  names(df) <- as.character(unlist(df[1,]))
  df[-1,]
}
transpose_header <- header.true(transpose_df)
# CHANGED filter for the countries we want to see
#new_df <- transpose_header %>% select(Argentina,Brazil,Chile,US)
# make the index the first row
d<-setDT(transpose_header, keep.rownames = TRUE)[]
# stack or melt the matrix
mdata <- melt(d, id=c("rn"))
# rename columns
df2 <- rename(mdata, date = rn,country = variable, confirmed = value )
# clean for dates
df2$date = sub('X', '', df2$date)
# set data types
df2$date = as.Date(df2$date, format='%m.%d.%y')
df2$confirmed = as.numeric(as.character((df2$confirmed)))
df2$country = as.character(df2$country)
# create a new column continent\
region2 <- rename(region, country = Country, continent = Continent)
full <- merge(df2, region2, by = c("country") , all.x = TRUE)
# create another new column for population
pop_df <- pop %>% select(country, X2020)
full_df <- merge(full, pop_df, by = c("country") , all.x = TRUE)
#######################
# add together by region
```


```{r, include=FALSE}
tot<-full_df %>% filter(date>="2020-03-27")
region_tot<-tot %>% group_by(continent) %>% summarize(totals=sum(confirmed)) %>% drop_na()
```

# LATAM has just `r format(round((region_tot[6,2]/sum(region_tot$totals))*100, 2), nsmall = 1)` percent of world wide cases.

```{r bar chart}  
ggplot(region_tot, aes(x=continent,y=totals)) + theme_bw()+ geom_bar(fill=rgb(0.1,0.4,0.5,0.7),stat="identity")+scale_y_continuous(labels =scales::comma_format())+labs(y = "confirmed cases", x = "continent",
title = "Global Confirmed COVID-19 Cases")
```
***

# Times series of continents  
## Regions of the third world continue to show low numbers of cases
```{r country filter region, include=FALSE}
region_time <-full_df %>% group_by(continent,date) %>% summarize(totals=sum(confirmed)) %>% drop_na()
```

```{r first_line, fig.width=8, fig.height=6}
ggplot(region_time, aes(date,totals, color=continent)) + geom_line(size=1) + theme_bw()+scale_y_continuous(labels =scales::comma_format())+ labs(y = "confirmed cases", x = "date", title = "Confirmed COVID-19 Cases by Continent")
```
***
# Regional confirmed cases as a ratio to population

```{r, include=FALSE}
# get the total population pero continent
tot_pop <- tot %>% group_by(continent) %>% summarize(totals=sum(as.numeric(X2020),na.rm = TRUE)) %>% drop_na()
tot_d <- merge(region_tot, tot_pop, by = c("continent") , all.x = TRUE)
tot_df <-rename(tot_d, confirmed = totals.x, population = totals.y)
tot_df <- tot_df %>% mutate(ratio_pop = confirmed/population*100)
```



```{r second_bar, fig.width=8, fig.height=6}
ggplot(tot_df, aes(x=continent,y=ratio_pop)) + theme_bw()+ geom_bar(fill=rgb(0.1,0.4,0.5,0.7),stat="identity")+labs(y = "percent of population infected", x = "continent",
title = "Percent of population infected by Continent")
```


