---
output:
  html_document:
    title: "COVIZ: COVID_19 in Latin America"
    author: "Brandon Janes"
    date: "3/22/2020"
    theme: cerulean
    highlight: haddock
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: no
    toc_depth: '3'
---
# COVIZ: COVID-19 in Latin America
By Brandon Janes
```{r import libraries,message=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(data.table)
```
# **Upload Data**
Data from repository for the 2019 Novel Coronavirus Visual Dashboard operated by the [**Johns Hopkins University Center for Systems Science and Engineering**](https://github.com/CSSEGISandData/COVID-19) (JHU CSSE). Also, countries and continents from [**dbouquin**](https://github.com/dbouquin/IS_608/blob/master/NanosatDB_munging/Countries-Continents.csv).
```{r, import data}
confirmed <- read.csv("./data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv", stringsAsFactors = FALSE)
region <- read.csv("./data/country-continent.csv", stringsAsFactors = FALSE)
```
```{r data wrangling, include=FALSE}
confirm <- select(confirmed, -(Lat:Long))
matriz <- select(confirm, -(Province.State))
matriz[-1] = as.numeric(unlist(matriz[-1]))
matriz[is.na(matriz)] <- 0
# group by
df <- matriz %>% group_by(Country.Region) %>% summarise_all(funs(sum))
# transpose
transpose_df <- as.data.frame(t(df))
# make country names column headers with custom function
header.true <- function(df) {
  names(df) <- as.character(unlist(df[1,]))
  df[-1,]
}
transpose_header <- header.true(transpose_df)
# CHANGED filter for the countries we want to see
#new_df <- transpose_header %>% select(Argentina,Brazil,Chile,US)
# make the index the first row
d<-setDT(transpose_header, keep.rownames = TRUE)[]
# stack or melt the matrix
mdata <- melt(d, id=c("rn"))
# rename columns
df2 <- rename(mdata, date = rn,country = variable, confirmed = value )
# clean for dates
df2$date = sub('X', '', df2$date)
# set data types
df2$date = as.Date(df2$date, format='%m.%d.%y')
df2$confirmed = as.numeric(as.character((df2$confirmed)))
df2$country = as.character(df2$country)
# create a new column continent\
region2 <- rename(region, country = Country, continent = Continent)
full_df <- merge(df2, region2, by = c("country") , all.x = TRUE)
#######################
# add together by region
```

```{r, include=FALSE}
tot<-full_df %>% filter(date>="2020-03-23")
region_tot<-tot %>% group_by(continent,date) %>% summarize(totals=sum(confirmed))
```

```{r country filter, include=FALSE}
# select the countries we want to show
df3 <- df2 %>% filter(str_detect(country,"China")|str_detect(country,"US")|str_detect(country,"Argentina")|str_detect(country,"Brazil")|str_detect(country,"Chile")|str_detect(country,"Chile")|str_detect(country,"Italy")|str_detect(country,"Mexico")|str_detect(country,"Peru")|str_detect(country,"Colombia"))
```

# LATAM has just `r format(round((region_tot[7,3]/sum(region_tot$totals))*100, 2), nsmall = 1)` percent of world wide cases.

```{r}  
ggplot(region_tot, aes(x=continent,y=totals)) + theme_bw()+ geom_bar(fill=rgb(0.1,0.4,0.5,0.7),stat="identity")+scale_y_continuous(labels =scales::comma_format())+labs(y = "confirmed cases", x = "continent",
title = "Global Confirmed COVID-19 Cases")
```
***

# LATAM compared to other countries 
## Latin American countries remain very low in comparison to other regions of the world
```{r first_line, fig.width=8, fig.height=6}
ggplot(df3, aes(date,confirmed, color=country)) + geom_line(size=1) + theme_bw()+ labs(y = "confirmed cases", x = "date", title = "Time Series of Confirmed COVID-19 Cases in the World")
```

***


# LATAM exponential growth in the month of March
```{r country filter LATAM, include=FALSE}
# select the countries we want to show
df4 <- df2 %>% filter(str_detect(country,"Argentina")|str_detect(country,"Brazil")|str_detect(country,"Chile")|str_detect(country,"Chile")|str_detect(country,"Mexico")|str_detect(country,"Peru")|str_detect(country,"Colombia")|str_detect(country,"Ecuador")|str_detect(country,"Venezuela")|str_detect(country,"Cuba"))

df5 <- df4 %>% filter(date>=as.Date("2020-03-01"))
march <- df5 %>% filter(date<as.Date("2020-03-23"))
```

```{r second_line, fig.width=8, fig.height=6}
ggplot(march, aes(date,confirmed, color=country)) + geom_line(size=1) + theme_bw()+ labs(y = "confirmed cases", x = "date", title = "March Confirmed COVID-19 Cases in LATAM")
```


```{r data wrangling march, include=FALSE}
latam <- transpose_header %>% select(Argentina,Brazil,Chile,Ecuador,Mexico,Colombia)
corona<-setDT(latam, keep.rownames = TRUE)[]
# clean and format date column
corona$rn = sub('X', '', corona$rn)
corona$rn = as.Date(corona$rn, format='%m.%d.%y')
# fitler of march
march <- corona %>% filter(rn>=as.Date("2020-03-01"))
march <- march %>% filter(rn<as.Date("2020-03-23"))
march_latam <- melt(march, id=c("rn"))
# rename columns
covid <- rename(march_latam, date = rn,country = variable, confirmed = value )
covid$confirmed = as.numeric(as.character((covid$confirmed)))
covid$country = as.character(covid$country)
```


```{r include=FALSE}
#ggplot(covid, aes(date,confirmed, color=country)) + geom_line() + theme_classic()+ labs(y = "confirmed cases", x = "date", title = "March Confirmed COVID-19 Cases in LATAM")

#ggplot(covid, aes(date,confirmed, color=country)) + geom_line(size=1, alpha=0.8) + theme_classic()+ labs(y = "confirmed cases", x = "date", title = "March Confirmed COVID-19 Cases in LATAM")

```

